Implementing Capon Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging

### 2012-11-15 HRUIonGPU/jrnl_article.tex

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Implementing Capon Adaptive Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

The final output, as shown in (\ref{eq:z_mv}) and (\ref{eq:z_mv2}), can be formed by either reducing the data down to length $L$, or increase the weight vector to length $M$. Since we have decided to have $L$ threads per output, the data is reduced to length $L$ by letting thread $r$ calculate 
\begin{align}
\breve{x}_{r}^{\Delta} = \sum_{k=r}^{r+N_L-1}x_{k}^{\Delta} \text{ for } r \in [0, L-1].
\end{align}
Each thread now computes $z_r = w_r^*x_r$, and finally one thread computes and outputs $z = \sum_{r=0}^{L-1} z_r$ to global memory. The last sum can also be performed using a binary reduction sum across several threads for increased speed.

\section{Beamspace Processing}\label{sec:bs}
Nilsen et al. \cite{Nilsen2009} have proposed to apply beamspace processing to further reduce the required computations. With beamspace we mean transforming the channel data from element space to a fan of beams covering the sector illuminated by transmit beams. The transformation is written as
\begin{align}
\vec{x}_{BS} = \mat{B}\vec{x}^{\Delta}.
\end{align} 
The steering vectors in $\mat{B}$, the so-called Butler matrix, determine each beam's direction in space. We can therefore easily remove beams where no interference is present by removing rows in $\mat{B}$. For a focused system like a cardiac ultrasound scanner, where the received signal is concentrated in a narrow band around broadside, there is a small percentage of the beams that contains almost all the energy. Nilsen et al. concluded that as little as three beams around broadside were adequate to produce results comparable to applying Capon beamforming in element space.

The beamspace version of Capon is derived by replacing the covariance matrix in (\ref{eq:w}) with the beamspace covariance matrix $\mat{\hat{R}}_{BS} = f(\vec{x}_{BS})$, where $\mat{B} \in \mathbb{C}^{N_b,M}$, and $N_b$ is the number of selected beams. Note that when subarray averaging is applied, the transformation must be done per subarray, yielding subarrays that no longer overlap in space. Hence, the amount of data is change from $M$ to $N_LN_b$ per sample.

(Go in detail about the implementation using sliding DFT).

(Go in detail about what kind of modifications we need to do to each of the three steps)

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Implementing Capon Beamforming on the GPU for Real Time Cardiac Ultrasound Imaging}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

The final output, as shown in (\ref{eq:z_mv}) and (\ref{eq:z_mv2}), can be formed by either reducing the data down to length $L$, or increase the weight vector to length $M$. Since we have decided to have $L$ threads per output, the data is reduced to length $L$ by letting thread $r$ calculate 
\begin{align}\label{eq:subsum}
\breve{x}_{r}^{\Delta} = \sum_{k=r}^{r+N_L-1}x_{k}^{\Delta} \text{ for } r \in [0, L-1].
\end{align}
Each thread now computes $z_r = w_r^*x_r$, and finally one thread computes and outputs $z = \sum_{r=0}^{L-1} z_r$ to global memory. The last sum can also be performed using a binary reduction sum across several threads for increased speed.

\section{Beamspace Processing}\label{sec:bs}
Nilsen et al. \cite{Nilsen2009} have proposed to apply beamspace processing to further reduce the required computations for the Capon beamformer. With beamspace we mean transforming the channel data from element space to a fan of beams covering the sector illuminated by transmit beams. The transformation is written as
\begin{align}
\vec{x}_{BS}[n] &= \mat{B}\vec{x}^{\Delta}[n], \ \ \ \mat{B} \in C^{M,M}\\
[\mat{B}]_{p,q} &= \frac{1}{\sqrt{M}}e^{-j 2 \pi p q / M}.
\end{align} 
The steering vectors in $\mat{B}$, the so-called Butler matrix, determine each beam's direction in space. We can therefore easily remove beams where no interference is present by removing rows from $\mat{B}$.

For a focused system like cardiac ultrasound imaging, where the received signal is concentrated in a narrow band around broadside, there is a small percentage of the beams that contains almost all the energy. Nilsen et al. concluded that as little as three beams around broadside were adequate to produce results comparable to applying Capon beamforming in element space. Hence, the inversion steps is found to be constant with respect to the number of array elements.

The beamspace version of the Capon beamformer is derived by replacing the covariance matrix in (\ref{eq:w}) with the beamspace covariance matrix $\mat{\hat{R}}_{BS} = f(\vec{x}_{BS}) = f(\mat{B}\vec{x})$, where $\mat{B} \in \mathbb{C}^{N_b,L}$, and $N_b$ is the number of selected beams. Note that when subarray averaging is applied, the transformation must be done per subarray, yielding subarrays that no longer overlap in space. If we want to apply the transformation before the covariance matrix is constructing, the amount of data is change from $M$ to $N_LN_b$ per sample. The implementation calculating covariance matrices needs to be adjusted to handle non-overlapping subarrays. BuildR complexity is changed to $O(N_b^2L)$, so constructing the covariance matrices will be faster in beamspace as long as $N_b < \sqrt{L}$. However it will involve large changes to the processing pipeline described in Section \ref{sec:meth}.

In both cases we need to solve using a transformed steering vector
\begin{align}
\vec{a}_{BS} &= \mat{B}\vec{1}_L = \sqrt{L}\vec{e}_1,
\end{align}
where $\vec{e}_i$ has value one in the i$^{th}$ position and is zero everywhere else.

Another choice is to transform the covariance matrix to beamspace. 
\begin{align}
\mat{\hat{R}}_{BS} &= \mat{B}\mat{\hat{R}}\mat{B}^H\\
\end{align}
By this approach we go from solving $\mat{\hat{R}}\vec{b} = \vec{1}_L$, which is an $L \times L$ system, to solving the $N_b \times N_b$ system $\mat{\hat{R}}_{BS}\vec{b}_{BS} = (\sqrt{L}\vec{e}_1)$. Now, one can either choose to transform the subarray sums, as in (\ref{eq:subsum}), to beamspace or transform the result, $\vec{b}_{BS}$, back to element space.

In our implementation we have tried to minimize the number of adjustments that had to be done in order to integrate beamspace processing into the GPU-Capon processing pipeline. The beamspace-results presented in this paper is therefore produced by transforming covariance matrices to beamspace, and after solving, transforming the result back to element space.

%(Go in detail about the implementation using sliding DFT).

%(Go in detail about what kind of modifications we need to do to each of the three steps)

%Note that $\mat{B}\mat{B}^H = \mat{I}$ and $\mat{B}^H\mat{B} \ne \mat{I}$ when $N_b < L$.

%Another way of computing the beamspace transformation is to transform the steering vector $\vec{a}$ and covariance vector $\mat{\hat{R}}$ beforing solving the following system of linear equations: $(\mat{B}\mat{\hat{R}}\mat{B}^H)^{-1}(\mat{B}a) = b_{BS}$.
